---
title: "R_aging_battleship_analyses"
author: "Abigail G Hedden (Shivers)"
date: "2024-08-01"
output: html_document
---

# Project Set-Up
## Set working directory and load required packages 
```{r setup, include=FALSE} 

knitr::opts_chunk$set(echo = TRUE)

# set working directory
# setwd("")

# set mirror for cran to allow for knitting
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)

# install and call required packages 
# if (!require("package")) {install.packages("package"); require("package")}

if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("plyr")) {install.packages("plyr"); require("plyr")}
if (!require("readr")) {install.packages("readr"); require("readr")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("tidytext")) {install.packages("tidytext"); require("tidytext")}
if (!require("textdata")) {install.packages("textdata"); require("textdata")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")} 
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("wordcloud2")) {install.packages("wordcloud2"); require("wordcloud2")} 
if (!require("wordcloud")) {install.packages("wordcloud"); require("wordcloud")} 
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")} 
if (!require("data.table")) {install.packages("data.table"); require("data.table")} 
if (!require("changepoint")) {install.packages("changepoint"); require("changepoint")} 
if (!require("tibble")) {install.packages("tibble"); require("tibble")} 
if (!require("EntropyExplorer")) {install.packages("EntropyExplorer"); require("EntropyExplorer")} 
if (!require("entropy")) {install.packages("entropy"); require("entropy")} 
if (!require("knitr")) {install.packages("knitr"); require("knitr")} 
if (!require("lme4")) {install.packages("lme4"); require("lme4")}
if (!require("sjPlot")) {install.packages('sjPlot'); require("sjPlot")}
if (!require("see")) {install.packages('see'); require("see")}
if (!require("brms")) {install.packages('brms'); require("brms")}
if (!require("devtools")) {install.packages("devtools"); require("devtools")}
if (!require("ggpubr")) {install.packages('ggpubr'); require("ggpubr")}
if (!require("lmerTest")) {install.packages('lmerTest'); require("lmerTest")}
if (!require("dlookr")) {install.packages('dlookr'); require("dlookr")}
if (!require("cowplot")) {install.packages('cowplot'); require("cowplot")}
if (!require("gridGraphics")) {install.packages('gridGraphics'); require("gridGraphics")}
if (!require("ggeffects")) {install.packages('ggeffects'); require("ggeffects")}
if (!require("ggpubr")) {install.packages('ggpubr'); require("ggpubr")}
if (!require("robustlmm")) {install.packages('robustlmm'); require("robustlmm")}
if (!require("MASS")) {install.packages('MASS'); require("MASS")}
if (!require("ggbeeswarm")) {install.packages("ggbeeswarm"); require("ggbeeswarm")}
if (!require("mediation")) {install.packages("mediation"); require("mediation")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("plotrix")) {install.packages("plotrix"); require("plotrix")}

```

## Load in data

### Load in cleaned data preprocessed separately in MATLAB and python
```{r}
# created below
# data for the information and reward mixed effects model:
info_rew_mixed_eff_data <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/info_rew_mixed_eff_data.csv")

# data with choice distance 
cart_dist_ent_df <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/cart_dist_ent_df.csv")

# dataframe with forager scores, information and reward values, and changepoint for each participant
forscore_cpt_data <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/forscore_cpt_data.csv")

# totals df
totals_df <- read.csv('~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/totals_df.csv')
  # can also serve as a lookup table for later analyses that required obtaining the cpt or age of participants
  # summary output of mousetracking data cleaning and wrangling. see `aging_bship_mousetracking_wrangling.ipynb`

genders_df <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/bship_genders.csv")
```

## Create age/cpt/gender look-up table df (includes all participants)
```{r}

df <- merge(df, participant_info[, c("subject_id","age")], by = "subject_id")

```


```

# Wrangling

### Load in and wrangle data generated and semi-processed in MATLAB
Original data from the Battleship task was pre-processed in MATLAB. This first set of chunks takes these data and wrangles them to the dataframe(s) that are used throughout these analyses

#### Create dataframe for mixed effects models 
The purpose of this data will be to compare how information, reward, trial number, and age impact the choice number using a mixed effects model.

This dataframe will hold Subject ID, Choice Number, Trial Number, Information, Reward, and age of all participants

##### Load in data from MATLAB
MATLAB data in csv format
```{r}
# load in MATLAB data from Box
options(scipen = 999) # prevents scientific notation in the table

# For each of the samples below:
# Given a .csv file (MATLAB output), load the csv contents into a dataframe in R and rename 3 of the columns
# Returns: 3 dataframes, one for each of the 3 samples. Each dataframe contains 5 columns (Choice Number, Information, Reward, Subject ID, and Trial Number). Row number varies dependent on subject variability in trial and choice counts across the task. 

# Sample 1 (ages 18-35)
#young_multilin_table <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/YA_multilevel_data.csv")
YA_mixeff <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/YA_mixeff_data.csv")
colnames(YA_mixeff)[1] = "Choice_Number"
colnames(YA_mixeff)[2] = "Information"
colnames(YA_mixeff)[3] = "Reward"

# Sample 2 (ages 35-50)
MA_mixeff <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/MA_mixeff_data.csv") #
colnames(MA_mixeff)[1] = "Choice_Number"
colnames(MA_mixeff)[2] = "Information"
colnames(MA_mixeff)[3] = "Reward"

# Sample 3 (ages 50-75)
OA_mixeff <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/OA_mixeff_data.csv") #
colnames(OA_mixeff)[1] = "Choice_Number"
colnames(OA_mixeff)[2] = "Information"
colnames(OA_mixeff)[3] = "Reward"
```

##### Add age information and merge to 1 combined dataframe
Given 3 dataframes, one for each sample, create 1 dataframe that adds age and merges the three to be used in various analyses

**Returns:**
all_mixed_eff : A dataframe with 6 columns (Choice Number, Information, Reward, Subject ID, Trial, and Age) and 88428 rows, each row being a unique choice in a unique trial for a unique subject. This data file shows the expected information and reward associated with the tile choice chosen in each unique choice, in each unique trial, by each unique subject. 

```{r}
# load in age information
# young adults (ages 18-35)
ya_ages_list <- c(20, 19, 19, 18, 20, 22, 21, 20, 19, 19, 20, 19, 20, 24, 20, 19, 21, 19, 21, 19, 20, 22, 20, 19, 20, 26, 19, 19, 18, 20, 20, 19, 21, 20, 18, 19, 20, 20, 22, 20, 20, 21, 22, 19, 19, 21, 19, 19, 20, 20, 20, 20, 21, 20, 20, 21, 21)

# middle adults (ages 35-50)
ma_ages_list <- c(44, 42, 42, 48, 35, 45, 39, 39, 39, 37, 40, 38, 38, 40, 39, 39, 43, 37, 46, 36, 39, 42, 45, 38, 35, 38, 38, 47, 37, 40, 41, 43, 48, 40, 41, 43, 36, 43, 43, 48, 43, 42, 38, 38, 48, 43, 38, 37, 40, 43, 36, 47, 44, 46, 39, 40, 45, 37, 39)

# older adults  (ages 50-75)
oa_ages_list <- c(54, 64, 51, 53, 71, 58, 55, 52, 52, 57, 51, 55, 52, 60, 62, 56, 56, 60, 57, 59, 63, 64, 61, 57, 59, 56, 56, 56, 59, 52, 56, 71, 55, 64, 50, 54, 56, 53, 65, 70, 54, 51, 66, 54, 50, 51, 57, 53, 56, 51, 52, 57, 51)


# define function to create a simple ages dataframe from a list
#' create an ages dataframe from a list of ages
#' 
#' @param ages_list a vector of ages
#' @param subject_count number of subjects
#' @return a dataframe with Subject and age columns
create_ages_df <- function(ages_list, subject_count) {
  tibble(
    Subject = 1:subject_count,
    Age = ages_list
  )
}

# define function to add ages to the mixed effects data
#' add age information to a mixed effects dataset
#' 
#' @param mixeff_data a dataframe containing choice data (Choice_Number, Information, Reward, Subject, Trial)
#' @param ages_df a dataframe with Subject and age columns 
#' @param id_prefix prefix to add to subject IDs (e.g., "YA", "MA", "OA")
#' @return a dataframe with age information added and subject IDs modified
add_age_to_dataset <- function(mixeff_data, ages_df, id_prefix) {
  # ensure column names are correct
  colnames(mixeff_data)[1:3] <- c("Choice_Number", "Information", "Reward")
  
  # join the age data to the mixed effects data
  result <- mixeff_data %>%
    left_join(ages_df, by = "Subject") %>%
    # add prefix to Subject ID
    mutate(Subject = paste0(id_prefix, Subject))
  
  return(result)
}

### load in ages and add to mixed-effects dataframe
# create age dataframes
ya_ages <- create_ages_df(ya_ages_list, length(ya_ages_list))
ma_ages <- create_ages_df(ma_ages_list, length(ma_ages_list))
oa_ages <- create_ages_df(oa_ages_list, length(oa_ages_list))

# add ages to datasets
YA_mixeff_table <- add_age_to_dataset(YA_mixeff, ya_ages, "YA")
MA_mixeff_table <- add_age_to_dataset(MA_mixeff, ma_ages, "MA")
OA_mixeff_table <- add_age_to_dataset(OA_mixeff, oa_ages, "OA")

# combine all datasets into one
all_mixed_eff <- bind_rows(YA_mixeff_table, MA_mixeff_table, OA_mixeff_table)

# save as csv
write.csv(all_mixed_eff, file = "info_rew_mixed_eff_data.csv", row.names = FALSE)
```





## Load in foraging and changepoint data
### Create a new dataframe that contains the average information and reward for each subject. 
```{r} 
tmp_avg_df <- data.frame(matrix(ncol = 4, nrow = 1)) # create temporary empty dataframe to be used to fill the for loop below
colnames(tmp_avg_df) <- c('subject', 'age', 'avg_info', 'avg_rew') # provide column names

# create an empty data frame that will be filled by for loop below
info_rew_means <- c()
                    
# make table with average information and average reward for each participant
for(s in unique(all_mixed_eff$Subject)){
  sub_mean = all_mixed_eff[all_mixed_eff$Subject==s,]
  tmp_avg_df$subject = s
  tmp_avg_df$age = mean(sub_mean$Age) # doing mean here which will just be their age since all age values are the same per subject
  tmp_avg_df$avg_info=mean(sub_mean$Information)
  tmp_avg_df$avg_rew=mean(sub_mean$Reward)
  info_rew_means = rbind(info_rew_means, tmp_avg_df)
}

```


### Load in Changepoints and Forager Scores
```{r}
# input values for mean final changepoints in each group --- these are all subjects changepoints and includes 0's DO NOT USE FOR ANY OTHER ANALYSES
all_YA_cpts = c(84, 33, 49, 67, 39, 66, 36, 68, 64, 11, 58, 36, 81, 24, 31, 26, 77, 0, 65, 84, 49, 24, 21, 51, 72, 12, 69, 60, 58, 24, 51, 0, 76, 22, 58, 36, 72, 52, 59, 53, 55, 22, 62, 30, 44, 51, 42, 9, 0, 66, 27, 67, 38, 40, 3, 49, 33)
all_MA_cpts <- c(51, 80, 29, 13, 43, 83, 35, 58, 55, 0, 31, 43, 33, 33, 41, 37, 11, 57, 34, 56, 63, 12, 13, 35, 60, 0, 58, 15, 11, 31, 33, 36, 31, 64, 29, 36, 46, 36, 55, 39, 0, 37, 8, 77, 0, 46, 55, 22, 67, 29, 53, 30, 16, 28, 38, 23, 33, 74, 50)
all_OA_cpts <- c(82, 26, 25, 64, 52, 26, 40, 0, 2, 17, 0, 60, 53, 25, 5, 44, 16, 50, 15, 24, 13, 0, 21, 32, 23, 16, 59, 48, 0, 63, 56, 45, 0, 62, 7, 53, 70, 57, 31, 0, 15, 30, 30, 44, 51, 68, 48, 73, 44, 76, 67, 35, 36)

all_cpts <- c(all_YA_cpts, all_MA_cpts, all_OA_cpts)

# load in forager scores for all groups 
# these scores have the 0's removed
# young -- 53/57
YA_info_for_s <- c(0.4, 0.68, 0.457142857, 0.38, 0.517647059, 0.555555556, 0.552941176, 0.614285714, 0.257142857, 0.866666667, 0.414285714, 0.74, 0.6, 0.58, 0.542857143, 0.4, 0.323076923, 0.453333333, 0.475, 0.4, 0.714285714, 0.82, 0.654545455, 0.705882353, 0.866666667, 0.481818182, 0.554545455, 0.493333333, 0.511111111, 0.581818182, 0.333333333, 0.85, 0.378947368, 0.825, 0.45, 0.530434783, 0.483333333, 0.563636364, 0.509090909, 0.7, 0.390909091, 0.52, 0.768421053, 0.529411765, 0.6, 1, 0.427272727, 0.581818182, 0.575, 0.7, 0.2, 0.5, 0.542857143)
YA_rew_for_s <- c(0.347368421, 0.24, 0.357142857, 0.38, 0.188235294, 0.488888889, 0.247058824, 0.3, 0.4, 0.333333333, 0.357142857, 0.26, 0.306666667, 0.26, 0.342857143, 0.32, 0.369230769, 0.493333333, 0.3625, 0.3, 0.342857143, 0.24, 0.327272727, 0.329411765, 0.333333333, 0.363636364, 0.327272727, 0.346666667, 0.222222222, 0.290909091, 0.371428571, 0.25, 0.347368421, 0.2, 0.36, 0.391304348, 0.366666667, 0.363636364, 0.309090909, 0.3, 0.372727273, 0.3, 0.315789474, 0.317647059, 0.288888889, 0.2, 0.372727273, 0.327272727, 0.325, 0.25, 0.36, 0.325, 0.542857143)

# middle age -- 55/59
MA_info_for_s <- c(0.72, 0.8, 0.742857143, 0.433333333, 0.709090909, 0.516666667, 0.707692308, 0.7, 0.588235294, 0.690909091, 0.7, 0.6, 0.666666667, 0.675, 0.646153846, 0.52, 0.552941176, 0.733333333, 0.533333333, 0.46, 0.733333333, 1, 0.771428571, 0.516666667, 0.444444444, 0.816666667, 0.4, 0.72, 0.784615385, 0.766666667, 0.825, 0.436363636, 0.75, 0.56, 0.6, 0.6, 0.6, 0.576923077, 0.4, 0.8, 0.411764706, 0.745454545, 0.4, 0.657142857, 0.484210526, 0.833333333, 0.583333333, 0.833333333, 0.35, 0.828571429, 0.682352941, 0.84, 0.48, 0.695652174, 0.753846154)
MA_rew_for_s <- c(0.32, 0.35, 0.228571429, 0.366666667, 0.290909091, 0.4, 0.292307692, 0.233333333, 0.282352941, 0.254545455, 0.275, 0.26, 0.3, 0.25, 0.261538462, 0.3, 0.423529412, 0.266666667, 0.4, 0.36, 0.2, 0.2, 0.228571429, 0.366666667, 0.366666667, 0.3, 0.371428571, 0.26, 0.230769231, 0.166666667, 0.2, 0.327272727, 0.25, 0.32, 0.386666667, 0.28, 0.307692308, 0.292307692, 0.4, 0.2, 0.364705882, 0.254545455, 0.333333333, 0.171428571, 0.410526316, 0.266666667, 0.2, 0.2, 0.425, 0.371428571, 0.223529412, 0.08, 0.333333333, 0.391304348, 0.246153846)
  
# old -- 46/53
OA_info_for_s <- c(0.372413793, 0.672727273, 0.727272727, 0.466666667, 0.625, 0.781818182, 0.769230769, 0.8, 0.55, 0.5, 0.885714286, 0.65, 0.75, 0.8, 0.75, 0.75, 0.8, 0.85, 0.6, 0.975, 0.628571429, 0.777777778, 0.222222222, 0.613333333, 0.488888889, 0.741176471, 0.65, 0.607407407, 0.75, 0.533333333, 0.541176471, 0.633333333, 0.685714286, 0.9, 0.618181818, 0.694117647, 0.644444444, 0.536842105, 0.367741935, 0.775, 0.327272727, 0.725, 0.523529412, 0.384615385, 0.72, 0.675)
OA_rew_for_s <- c(0.351724138, 0.272727273, 0.236363636, 0.333333333, 0.275, 0.2, 0.246153846, 0.2, 0.35, 0.322222222, 0.257142857, 0.35, 0.233333333, 0.32, 0.25, 0.15, 0.225, 0.25, 0.2, 0.225, 0.2, 0.155555556, 0.377777778, 0.28, 0.266666667, 0.317647059, 0.28, 0.325925926, 0.35, 0.255555556, 0.282352941, 0.333333333, 0.228571429, 0.233333333, 0.309090909, 0.282352941, 0.311111111, 0.410526316, 0.4, 0.2375, 0.390909091, 0.2, 0.358823529, 0.415384615, 0.213333333, 0.375)

# make list of all info forager scores and all reward forager scores
all_info_fs <- c(YA_info_for_s, MA_info_for_s, OA_info_for_s)
all_rew_fs <- c(YA_rew_for_s, MA_rew_for_s, OA_rew_for_s)

# make new dataframe to allow us to remove rows where there are no changepoints/forager scores
# start out with the large dataframe that includes the average information and reward for each participant
info_rew_means_cpt_fs <- info_rew_means

# add in all changepoints, even those with 0's 
info_rew_means_cpt_fs$changepoint <- all_cpts

# remove people who did not get a forager score/changepoint
# we cannot just remove the rows where changepoint == 0  because there are some participants who had a changepoint but not a forager score, so we are going in here and just specifically removing the subjects who don't have a changepoint and/or a forager score
# YA 18, 32, 49, 55
# MA 10, 26, 41, 45
# OA 8, 9, 11, 22, 29, 33, 40
# 15 participants total

info_rew_means_cpt_fs <- info_rew_means_cpt_fs[!(info_rew_means_cpt_fs$subject %in% c("YA18", "YA32", "YA49", "YA55", "MA10", "MA26", "MA41", "MA45", "OA8", "OA9", "OA11", "OA22", "OA29", "OA33", "OA40")),]

# now we have removed all rows from our dataframe for the subjects who did not have a changepoint/forager score(s). now let's add in the forager scores to this dataframe to allow for analyses

# add forager scores to mean tables
info_rew_means_cpt_fs$info_fs <- all_info_fs
info_rew_means_cpt_fs$rew_fs <- all_rew_fs

# write.csv(info_rew_means_cpt_fs, file = "forscore_cpt_data.csv", row.names = FALSE)
```


# Mixed Effects Model for Information and Reward Utilization/Prioritization
Investigate the effect of information, reward, trial number, and age on choice number

Run a mixed-effects linear regression to regress **choice number** in trial against **trial number** in session, the expected **reward** for the chosen tile, the expected **information** for the chosen tile, and **age**. Also, test for all possible interactions. Include intercepts for subjects for random effects. 

**Parameters:**

- Choice_Number: The choice number in a given trial, where a choice is a selection of a tile on the grid
- Information: The expected information for the chosen tile on the given choice, where information is the difference in the entropy, or uncertainty, before and after each choice outcome
- Reward: The expected reward for the chosen tile on the given choice, where reward for a given tile is divided by the total number of times that tile had been previously selected
- Trial: The trial number at which a choice was made
- Age: The age of the participant
- Subject: The unique Subject ID for each participant

**Results:**

- As choice number increased, the use of information on tile selection decreased 
- The use of reward on tile selection increased with increasing choice number
- A significant interaction between information and reward was found 
- A significant interaction was also found between trial number and age 
- A three-way interaction was also found between information, reward, and trial number 
- A significant three-way interaction between information, trial number, and age
```{r}
info_rew_mix_mod <- lmerTest::lmer(Choice_Number ~ 1 + Information*Reward*Trial*Age + (1|Subject), data=info_rew_mixed_eff_data, control=lmerControl(optCtrl=list(optimizer='Nelder-Mead',maxfun=2e8)))
summary(info_rew_mix_mod)
```

# Information and Reward Analyses
## Information and Reward Variables
Run a mixed-effects linear regression to regress **information** against **reward**. Include intercepts for subjects for random effects. 


**Parameters:**

- Information: The expected information for the chosen tile on the given choice, where information is the difference in the entropy, or uncertainty, before and after each choice outcome
- Reward: The expected reward for the chosen tile on the given choice, where reward for a given tile is the number of of times choosing that tile revealed a portion of the shape (hits) in the past divided by the total number of times that tile had been previously selected
- Subject: The unique Subject ID for each participant
             

**Results:**

- Information and reward are negatively correlated (β = -0.021 ± 0.006, p < 0.001)
```{r}
rew_info_rel_model=lmer(Information ~ Reward + (Reward|Subject), data=info_rew_mixed_eff_data)
summary(rew_info_rel_model)
```

## Information and Reward Relationship as Trials Progress
Run a mixed-effects linear regression to regress **information** against **reward** and **trial number** as well as the **reward X trial number** interaction. Include intercepts for subjects for random effects. 


**Parameters:**

- Information: The expected information for the chosen tile on the given choice, where information is the difference in the entropy, or uncertainty, before and after each choice outcome
- Reward: The expected reward for the chosen tile on the given choice, where reward for a given tile is the number of of times choosing that tile revealed a portion of the shape (hits) in the past divided by the total number of times that tile had been previously selected
- Trial: The trial number at which a choice was made
- Subject: The unique Subject ID for each participant


**Results:**

- Information and reward become more negatively correlated as trial number increases (β = -0.003 ± 0.0003, p < 0.0001)
```{r}
rew_info_trial_rel_model=lmer(Information ~ Reward*Trial + (Reward*Trial|Subject), data=info_rew_mixed_eff_data)
summary(rew_info_trial_rel_model)
```


# Choice Distance Mixed Effects Model
Investigate the effect of age, choice number, and trial number on pairwise choice distance. 

Run a mixed-effects linear regression to regress **choice distance** between each pair of choices in trial against **age** of the participant, **choice number** in trial, and **trial number** in session. Also, test for all possible interactions. Include intercepts for subjects for random effects. 


**Parameters:**

- choice_dist: The choice number in a given trial, where a choice is a selection of a tile on the grid
- age: The age of the participant
- choice_num: The choice number in a given trial, where a choice is a selection of a tile on the grid
- trial_num: The trial number at which a choice was made
- subject_id: The unique Subject ID for each participant

**Results:**

- As age increased, the distance between pairs of choices also increased
- The distance between choices grew as trials progressed 
- Significant interactions between the following:
  - Age and choice number
  - Age and trial number
  - Choice number and trial number
- A three-way interaction between age, choice number, and trial number reveals that the effect of choice number and trial number on choice distance is weaker in older adults
```{r}
choice_dist_mix_mod <- lmerTest::lmer(choice_dist ~ age*choice_num*trial_num + (1|subject_id), data=cart_dist_ent_df) 
summary(choice_dist_mix_mod)
```


# Aging and Learning Speed (i.e., final changepoint)
Run a linear regression to regress **changepoint** against **age** of the participant.

**Parameters:**

- changepoint: The trial number at which the changepoint detection test detected learning of the shape set. Lower changepoint indicates a sooner trial and thus indicates faster learning.
- age: The age of the participant
        
**Results:**

- With increasing age, the final changepoint trial occurs earlier (β = -0.267 ± 0.103, p = 0.01)
- Older adults learn the shapes more quickly
```{r}
age_cpt_mod <- lm(changepoint ~ age, data = forscore_cpt_data)
summary(age_cpt_mod)
```

# General Task Completion Analyses
## Total Trials Completed and Age
Run a linear regression to regress **total trials** in session against **age** of the participant.

**Parameters:**

- tot_trials: The total number of trials completed by the participant during the task
- age: The age of the participant

**Results:**

- As age increased, the number of trials completed in the task decreased (β = -0.199 ± 0.084, p = 0.018)
- Older age is associated with completing fewer trials
```{r}
age_trial_corr_model <- lm(tot_trials ~ age, data = totals_df)
summary(age_trial_corr_model)
```

## Total Choices Made and Age
Run a linear regression to regress **total choices** in session against **age** of the participant.

**Parameters:**

- tot_choices: The **total** number of choices made by the participant in the task
- age: The age of the participant

**Results:**

- There was no relationship between the total number of choices made in the task and age
- Older age is not associated with more choices overall
```{r}
age_tot_choice_model <- lm(tot_choices ~ age, data = totals_df)
summary(age_tot_choice_model)
```

## Total Choices per Trial and Age
Run a linear regression to regress the **average choices** made against **age** of the participant.

**Parameters:**

- avg_choices: The average number of choices per trial for a given participant
- age: The age of the participant

**Results:**

- There is a positive relationship between the average number of choices per trial and age
- Older age is associated with making more choices per trial (β = 0.030 ± 0.007, p < 0.001)
```{r}
age_avg_choice_model <- lm(avg_choices ~ age, data = totals_df)
summary(age_avg_choice_model)
```

## Total Information per Trial and Age
Run a linear regression to regress the **average amount of information per trial** against **age** of the participant

**Parameters:**

- t_prop_tot_info: The average amount of information gain per trial for a given participant
- age: The age of the participant

**Results:**

- As age increased, the average amount of information per trial increased (β = 0.023 ± 0.005, p < 0.001)
- Older age is associated with gaining more information per trial
```{r}
age_avg_info_model <- lm(t_prop_tot_info ~ age, data = totals_df)
summary(age_avg_info_model)
```

# Foraging Abilities and Learning Analyses
## Information Forager Score and Reward Forager Score 
Run Pearson correlation test to test the relationship between **information forager score** and **reward forager score**

**Parameters:**

- info_fs: The information forager score
- rew_fs: The reward forager score

**Results:**

-  Information forager scores and reward forager scores are significantly anticorrelated (r = -0.629, p < 0.001) 
```{r}
cor.test(forscore_cpt_data$info_fs, forscore_cpt_data$rew_fs, method = "pearson")
```

## Age and Information Forager Score
Run a linear regression to regress **information forager score** against **age** of participant

**Parameters:**

- info_fs: The information forager score for a given participant
- age: The age of the participant

**Results:**

- Older age is associated with higher information forager scores (β = 0.003 ± 0.0008, p < 0.001) 
```{r}
age_info_fs_mod <- lm(info_fs ~ age, data = forscore_cpt_data)
summary(age_info_fs_mod)
```

## Age and Reward Forager Score
Run a linear regression to regress **reward forager score** against **age** of the participant

**Parameters:**

- rew_fs: The reward forager score for a given participant
- age: The age of the participant

**Results:**

- Older age is associated with lower reward forager scores (β = -0.001 ± 0.0004, p < 0.001).
```{r}
age_rew_fs_mod <- lm(rew_fs ~ age, data = forscore_cpt_data)
summary(age_rew_fs_mod)
```

## Information Forager Score and Changepoint
Run a linear regression to regress **changepoint** against **information forager score** and **age** of the participant, along with the **information forager score X age** interaction. 

**Parameters:**

- changepoint: The trial number at which the changepoint detection test detected learning of the shape set. Lower changepoint indicates a sooner trial and thus indicates faster learning
- info_fs: The information forager score for a given participant
- age: The age of the participant

**Results:**

- The ability to forage for information was associated with a lower changepoint (β = -63.149 ± 22.569, p = 0.006)
- In other words, the ability to forage for information is associated with faster learning
- There was no main effect of age or interaction between age and information forager score
```{r}
info_fs_cpt_age <- lm(changepoint ~ info_fs + age + info_fs:age, data=forscore_cpt_data)
summary(info_fs_cpt_age)
```

## Reward Forager Score and Changepoint
Run a linear regression to regress **changepoint** against **reward forager score** and **age** of the participant, along with the **reward forager score X age** interaction. 

**Parameters:**

- changepoint: The trial number at which the changepoint detection test detected learning of the shape set. Lower changepoint indicates a sooner trial and thus indicates faster learning
- rew_fs: The reward forager score for a given participant
- age: The age of the participant

**Results:**

- The ability to forage for reward was associated with a higher changepoint (β = 127.684 ± 56.382, p = 0.025)
- In other words, the ability to forage for reward is associated with slower learning
- There was no main effect of age or interaction between age and reward forager score
```{r}
rew_fs_cpt_age <- lm(changepoint ~ rew_fs + age + rew_fs:age, data=forscore_cpt_data)
summary(rew_fs_cpt_age)
```









