---
title: "cleaned_R_aging_battleship_analyses"
output: html_document
date: "2024-08-01"
---

# Load in Data and Required Packages
```{r setup, include=TRUE} 
knitr::opts_chunk$set(echo = TRUE)

# set mirror for cran to allow for knitting
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)

# install and call required packages 
# if (!require("package")) {install.packages("package"); require("package")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("knitr")) {install.packages("knitr"); require("knitr")} 
if (!require("lme4")) {install.packages("lme4"); require("lme4")}
if (!require("lmerTest")) {install.packages('lmerTest'); require("lmerTest")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")} 


if (!require("readr")) {install.packages("readr"); require("readr")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("tidytext")) {install.packages("tidytext"); require("tidytext")}
if (!require("textdata")) {install.packages("textdata"); require("textdata")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("wordcloud2")) {install.packages("wordcloud2"); require("wordcloud2")} 
if (!require("wordcloud")) {install.packages("wordcloud"); require("wordcloud")} 
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")} 
if (!require("data.table")) {install.packages("data.table"); require("data.table")} 
if (!require("changepoint")) {install.packages("changepoint"); require("changepoint")} 
if (!require("tibble")) {install.packages("tibble"); require("tibble")} 
if (!require("EntropyExplorer")) {install.packages("EntropyExplorer"); require("EntropyExplorer")} 
if (!require("entropy")) {install.packages("entropy"); require("entropy")} 
if (!require("sjPlot")) {install.packages('sjPlot'); require("sjPlot")}
if (!require("see")) {install.packages('see'); require("see")}
if (!require("brms")) {install.packages('brms'); require("brms")}
if (!require("devtools")) {install.packages("devtools"); require("devtools")}
if (!require("ggpubr")) {install.packages('ggpubr'); require("ggpubr")}
if (!require("dlookr")) {install.packages('dlookr'); require("dlookr")}
if (!require("cowplot")) {install.packages('cowplot'); require("cowplot")}
if (!require("gridGraphics")) {install.packages('gridGraphics'); require("gridGraphics")}
if (!require("ggeffects")) {install.packages('ggeffects'); require("ggeffects")}
if (!require("ggpubr")) {install.packages('ggpubr'); require("ggpubr")}
if (!require("robustlmm")) {install.packages('robustlmm'); require("robustlmm")}
if (!require("MASS")) {install.packages('MASS'); require("MASS")}
```

# Load in Data
```{r}
# data for the information and reward mixed effects model:
info_rew_mixed_eff_data <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/info_rew_mixed_eff_data.csv")

# data with Euclidean choice distance calculations
cart_dist_ent_df <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/cart_dist_ent_df.csv")

# dataframe with forager scores, information and reward values, and changepoint for each participant
forscore_cpt_data <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/forscore_cpt_data.csv")
  # excluded participants who did not get a forager score/changepoint
    # YA 18, 32, 49, 55
    # MA 10, 26, 41, 45
    # OA 8, 9, 11, 22, 29, 33, 40
    # 15 participants total excluded from this df and corresponding analyses

# df containing maximum scores earned by participants and total trials completed in the task
merged_scores_df <- read.csv('~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/merged_scores_df.csv')

# totals df: summary table of multiple task completion stats from mousetracking data
totals_df <- read.csv('~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/totals_df.csv')
  # can also serve as a lookup table for later analyses that required obtaining the cpt or age of participants
  # summary output of mousetracking data cleaning and wrangling. see `aging_bship_mousetracking_wrangling.ipynb`

# genders df: contains gender info for participants
genders_df <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/dataframes/bship_genders.csv")

# data with all mean directed exploration types calculated
mean_expl_df <- read.csv("~/Library/CloudStorage/Box-Box/Bakkour-Lab/projects/Battleship_task/analysis/mean_expl_df.csv")
```

# General Task Completion Analyses
This set of analyses examines general task completion variables, such as total trials completed and points earned, for example. We employed a variety of OLS linear regressions to investigate how these factors changed with age.

## Total Score and Age
Run a linear regression to regress the number of points earned, the **total score**, in session against **age** of the participant.

**Parameters:**

* `max_score`: The final score of the participant in the task
* `age`: The age of the participant

**Results:**
* As our participants’ age increases,  they earn fewer points overall by the end of the task (β = -0.961, 95% CI [-1.723, -0.200], p = 0.01).
* Older age is associated with lower overall scores.
```{r}
age_score_mod <- lm(max_score ~ age, data = merged_scores_df)
summary(age_score_mod)
confint(age_score_mod)
```


## Age and Learning Speed (i.e., final changepoint)
Run a linear regression to regress **changepoint** against **age** of the participant.

**Parameters:**
* `changepoint`: The trial number at which the changepoint detection test detected learning of the shape set. Lower changepoint indicates a sooner trial and thus indicates faster learning.
* `age`: The age of the participant
        
**Results:**
* With increasing age, the final changepoint trial occurs earlier (β = -0.267, 95% CI [-0.471, -0.064], p = 0.01)
* Older adults learn the shapes more quickly
```{r}
age_cpt_mod <- lm(changepoint ~ age, data = forscore_cpt_data)
summary(age_cpt_mod)
confint(age_cpt_mod)
```


## Total Trials Completed and Age
Run a linear regression to regress **total trials** in session against **age** of the participant.

**Parameters:**
* `tot_trials`: The total number of trials completed by the participant during the task
* `age`: The age of the participant

**Results:**
* As age increased, the number of trials completed in the task decreased (β = -0.199, 95% CI [-0.365, -0.035], p = 0.018)
* Older age is associated with completing fewer trials
```{r}
age_trial_corr_model <- lm(tot_trials ~ age, data = totals_df)
summary(age_trial_corr_model)
confint(age_trial_corr_model)
```


## Total Choices Made and Age
Run a linear regression to regress **total choices** in session against **age** of the participant.

**Parameters:**
* `tot_choices`: The **total** number of choices made by the participant in the task
* `age`: The age of the participant

**Results:**
* There was no relationship between the total number of choices made in the task and age
* Older age is not associated with more choices overall
```{r}
age_tot_choice_model <- lm(tot_choices ~ age, data = totals_df)
summary(age_tot_choice_model)
confint(age_trial_corr_model)
```


## Total Choices per Trial and Age
Run a linear regression to regress the **average choices** made against **age** of the participant.

**Parameters:**
* `avg_choices`: The average number of choices per trial for a given participant
* `age`: The age of the participant

**Results:**
* There is a positive relationship between the average number of choices per trial and age
* Older age is associated with making more choices per trial (β = 0.030, 95% CI [0.017, 0.044], p < 0.001).
```{r}
age_avg_choice_model <- lm(avg_choices ~ age, data = totals_df)
summary(age_avg_choice_model)
confint(age_avg_choice_model)
```

## Average Information per Trial and Age
Run a linear regression to regress the **average amount of information per trial** against **age** of the participant

**Parameters:**
* `t_prop_tot_info`: The average amount of information gain per trial for a given participant
* `age`: The age of the participant

**Results:**
* As age increased, the average amount of information per trial increased (β = 0.023, 95% CI [0.012, 0.033], p < 0.001).
* Older age is associated with gaining more information per trial
```{r}
age_avg_info_model <- lm(t_prop_tot_info ~ age, data = totals_df)
summary(age_avg_info_model)
confint(age_avg_info_model)
```

# Choice Distance Mixed Effects Model
Investigate the effect of age, choice number, and trial number on pairwise choice distance. 

Run a mixed-effects linear regression to regress **choice distance** between each pair of choices in trial against **age** of the participant, **choice number** in trial, and **trial number** in session. Also, test for all possible interactions. Include intercepts for subjects for random effects. 

**Parameters:**
* `choice_dist`: The choice number in a given trial, where a choice is a selection of a tile on the grid
* `age`: The age of the participant
* `choice_num`: The choice number in a given trial, where a choice is a selection of a tile on the grid
* `trial_num`: The trial number at which a choice was made
* `subject_id`: The unique Subject ID for each participant

**Results:**
* As age increased, the distance between pairs of choices also increased
* The distance between choices grew as trials progressed 
* Significant interactions between the following:
  + Age and choice number
  + Age and trial number
  + Choice number and trial number
* A three-way interaction between age, choice number, and trial number reveals that the effect of choice number and trial number on choice distance is weaker in older adults
```{r}
choice_dist_mix_mod <- lmerTest::lmer(choice_dist ~ age*choice_num*trial_num + (1|subject_id), data=cart_dist_ent_df) 
summary(choice_dist_mix_mod)
confint(choice_dist_mix_mod)
```

# Directed Exploration Analyses
Run a linear regression to regress **mean directed exploration** against **age** of the participant.

**Parameters:**
* `overall_dir_expl`: The mean weighted (by reward and information) directed exploration across trials
* `age`: The age of the participant
        
**Results:**
* Older age was associated with larger deviations from the expected distance, indicating more directed exploration (β = 8.136 ✕ 10-4, 95% CI [2.182 ✕ 10-4, 1.409 ✕ 10-3], p = 0.008).
```{r}
dir_expl_age_model <- lm(overall_dir_expl ~ age, data = mean_expl_df)
summary(dir_expl_age_model)
confint(dir_expl_age_model)
```

# Information and Reward Analyses
## Mixed Effects Model for Information and Reward Utilization/Prioritization
Investigate the effect of information, reward, trial number, and age on choice number.

Run a mixed-effects linear regression to regress **choice number** in trial against **trial number** in session, the expected **reward** for the chosen tile, the expected **information** for the chosen tile, and **age**. Also, test for all possible interactions. Include intercepts for participants for random effects. 

**Parameters:**
* `Choice_Number`: The choice number in a given trial, where a choice is a selection of a tile on the grid
* `Information`: The expected information for the chosen tile on the given choice, where information is the difference in the entropy, or uncertainty, before and after each choice outcome
* `Reward`: The expected reward for the chosen tile on the given choice, where reward for a given tile is divided by the total number of times that tile had been previously selected
* `Trial`: The trial number at which a choice was made
* `Age`: The age of the participant
* `Subject`: The unique Subject ID for each participant

**Results:**
* As choice number increased, the use of information on tile selection decreased 
* The use of reward on tile selection increased with increasing choice number
* A significant interaction between information and reward was found 
* A significant interaction was also found between trial number and age 
* A three-way interaction was also found between information, reward, and trial number 
* A significant three-way interaction between information, trial number, and age
```{r}
info_rew_mix_mod <- lmerTest::lmer(Choice_Number ~ 1 + Information*Reward*Trial*Age + (1|Subject), data=info_rew_mixed_eff_data, control=lmerControl(optCtrl=list(optimizer='Nelder-Mead',maxfun=2e8)))
summary(info_rew_mix_mod)
confint(info_rew_mix_mod)
```

## Relationship Beween Information and Reward
Run a mixed-effects linear regression to regress **information** against **reward**. Include intercepts for subjects for random effects. 

**Parameters:**
* `Information`: The expected information for the chosen tile on the given choice, where information is the difference in the entropy, or uncertainty, before and after each choice outcome
* `Reward`: The expected reward for the chosen tile on the given choice, where reward for a given tile is the number of of times choosing that tile revealed a portion of the shape (hits) in the past divided by the total number of times that tile had been previously selected
* `Subject`: The unique Subject ID for each participant
             
**Results:**
* Information and reward are negatively correlated (β = -0.021, 95% CI [-0.032, -0.009], p < 0.001).
```{r}
rew_info_rel_model=lmer(Information ~ Reward + (Reward|Subject), data=info_rew_mixed_eff_data)
summary(rew_info_rel_model)
confint(rew_info_rel_model)
```

## Information and Reward Relationship as Trials Progress
Run a mixed-effects linear regression to regress **information** against **reward** and **trial number** as well as the **reward X trial number** interaction. Include intercepts for subjects for random effects. 

**Parameters:**
* `Information`: The expected information for the chosen tile on the given choice, where information is the difference in the entropy, or uncertainty, before and after each choice outcome
* `Reward`: The expected reward for the chosen tile on the given choice, where reward for a given tile is the number of of times choosing that tile revealed a portion of the shape (hits) in the past divided by the total number of times that tile had been previously selected
* `Trial`: The trial number at which a choice was made
* `Subject`: The unique Subject ID for each participant

**Results:**
* Information and reward become more negatively correlated as trial number increases (β = -3.069 x 10-3, 95% CI [-3.762 x 10-3, -2.375 x 10-3], p < 0.0001)
```{r}
rew_info_trial_rel_model=lmer(Information ~ Reward*Trial + (Reward*Trial|Subject), data=info_rew_mixed_eff_data)
summary(rew_info_trial_rel_model)
confint(rew_info_trial_rel_model)
```


# Foraging Abilities and Learning Analyses
## Information Forager Score and Reward Forager Score 
Run Pearson correlation test to test the relationship between **information forager score** and **reward forager score**

**Parameters:**
* `info_fs`: The information forager score
* `rew_fs`: The reward forager score

**Results:**
*  Information forager scores and reward forager scores are significantly anticorrelated (r = -0.629, p < 0.001) 
```{r}
cor.test(forscore_cpt_data$info_fs, forscore_cpt_data$rew_fs, method = "pearson")
```


## Age and Information Forager Score
Run a linear regression to regress **information forager score** against **age** of participant

**Parameters:**
* `info_fs`: The information forager score for a given participant
* `age`: The age of the participant

**Results:**
* Older age is associated with higher information forager scores (β = 0.003 ± 0.0008, p < 0.001) 
```{r}
age_info_fs_mod <- lm(info_fs ~ age, data = forscore_cpt_data)
summary(age_info_fs_mod)
```


## Age and Reward Forager Score
Run a linear regression to regress **reward forager score** against **age** of the participant

**Parameters:**
* `rew_fs`: The reward forager score for a given participant
* `age`: The age of the participant

**Results:**
* Older age is associated with lower reward forager scores (β = -0.001 ± 0.0004, p < 0.001).
```{r}
age_rew_fs_mod <- lm(rew_fs ~ age, data = forscore_cpt_data)
summary(age_rew_fs_mod)
```


## Information Forager Score and Final Changepoint
Run a linear regression to regress **changepoint** against **information forager score** and **age** of the participant, along with the **information forager score X age** interaction. 

**Parameters:**
* `changepoint`: The trial number at which the changepoint detection test detected learning of the shape set. Lower changepoint indicates a sooner trial and thus indicates faster learning
* `info_fs`: The information forager score for a given participant
* `age`: The age of the participant

**Results:**
* The ability to forage for information was associated with a lower changepoint (β = -63.149, 95% CI [-107.744, -18.553], p = 0.006)
* In other words, the ability to forage for information is associated with faster learning
* There was no main effect of age or interaction between age and information forager score
```{r}
info_fs_cpt_age <- lm(changepoint ~ info_fs + age + info_fs:age, data=forscore_cpt_data)
summary(info_fs_cpt_age)
confint(info_fs_cpt_age)
```


## Reward Forager Score and Final Changepoint
Run a linear regression to regress **changepoint** against **reward forager score** and **age** of the participant, along with the **reward forager score X age** interaction. 

**Parameters:**
* `changepoint`: The trial number at which the changepoint detection test detected learning of the shape set. Lower changepoint indicates a sooner trial and thus indicates faster learning
* `rew_fs`: The reward forager score for a given participant
* `age`: The age of the participant

**Results:**
* The ability to forage for reward was associated with a higher changepoint (β = 127.684, 95% CI [16.278, 239.089], p = 0.025)
* In other words, the ability to forage for reward is associated with slower learning
* There was no main effect of age or interaction between age and reward forager score
```{r}
rew_fs_cpt_age <- lm(changepoint ~ rew_fs + age + rew_fs:age, data=forscore_cpt_data)
summary(rew_fs_cpt_age)
confint(rew_fs_cpt_age)
```


# Mediation Analysis
Run a mediation analysis to investigate whether information forager score mediates the relationship between age and final changepoint, using the “mediation” R software package (Tingley et al., 2014).

## Mediator Model: Regress the mediator on the independent variable.
**Parameters:**
* `info_fs`: The information forager score for a given participant (mediator)
* `age`: The age of the participant (independent variable)

**Results:**
* Age is positively associated with the information forager score
```{r}
mediator_mod <- lm(info_fs ~ age, data = info_rew_means_cpt_fs)
print(summary(mediator_mod))
confint(mediator_mod)
```

## Outcome Model: Regress the dependent variable on both the independent variable and the mediator.
**Parameters:**
* `changepoint`: The trial number at which the changepoint detection test detected learning of the shape set. Lower changepoint indicates a sooner trial and thus indicates faster learning (dependent variable)
* `info_fs`: The information forager score for a given participant (mediator)
* `age`: The age of the participant (independent variable)

**Results:**
* The direct effect of age on changepoint is negative but *not* statistically significant.
* Information forager score is strongly and negatively associated with changepoint. In other words, higher information forager scores are linked to lower changepoints (faster learning).
```{r}
# outcome model
outcome_mod <- lm(changepoint ~ age + info_fs, data = info_rew_means_cpt_fs)
print(summary(outcome_mod))
confint(outcome_mod)
```

## Mediation Analysis Model

**Parameters:**
* `model.m`: The **mediator model** (`mediator_mod`) predicting info_fs from age
* `model.y`: The **outcome model** (`outcome_mod`) predicting changepoint from age and info_fs
* `treat`: Treatment variable (`age`)
* `mediator`: Mediator variable (`info_fs`)
* `boot`: Logical flag indicating whether to use bootstrapping for confidence intervals (`TRUE`)
* `sims`: Number of bootstrap simulations (`10000`)

**Path Relationships Estimated:**
* **Direct effect (`ADE`):** Effect of age on changepoint not explained by info_fs
* **Indirect effect (`ACME`):** Effect of age on changepoint transmitted through info_fs
* **Total effect:** Combined direct and indirect effects of age on changepoint

**Mediation Analysis Results:**
* The total effect of age on final changepoint was significant, with increasing age associated with a lower final changepoint, as previously discussed. See `Total Effect` in below analyses.
* Mediation analysis revealed a significant indirect effect of age on changepoint through information forager score (β = -0.201, 95% CI [-0.3163, -0.09], p < 0.001), indicating that information forager score, at least partially, explains the relationship between age and final changepoint. See `ACME` in below analyses.
* After controlling for information forager score, the direct effect of age on final changepoint was attenuated, consistent with a complete mediation. See `ADE` in below analyses.
* Information forager score mediated approximately 75.25% of the total effect (95% CI [39.51, 203], p = 0.003), suggesting that a substantial portion of the association between age and changepoint operates through differences in information-seeking strategies. See `Prop. Mediated` in below analyses.
```{r}
info_age_fs_med <- mediate(
  model.m = mediator_mod,
  model.y = outcome_mod,
  treat = "age",
  mediator = "info_fs",
  boot = TRUE, # bootstrapping for more robust CIs
  sims = 10000 # number of simulations for bootstrap
)
# sample size used: 154 
print(summary(info_age_fs_med))
#               Estimate 95% CI Lower 95% CI Upper             p-value    
#ACME            -0.2012      -0.3163        -0.09 <0.0000000000000002 ***
#ADE             -0.0662      -0.2373         0.09              0.4428    
#Total Effect    -0.2674      -0.4664        -0.08              0.0028 ** 
#Prop. Mediated   0.7525       0.3951         2.03              0.0028 ** 



```





finishing up ...
- remove my working directory from the top, users will fill this in themselves
- remove path from loading in files\
